---
title: "Data Manipulation"
output: html_document
---

```{r setup, message=FALSE}
  library(dplyr)
  library(tidyverse)
```

#### **Load in the Dataset**
```{r, message=FALSE}
  litters_df = read_csv("../../Datasets/FAS_litters.csv")
  litters_df = janitor::clean_names(litters_df)

```

#### **Dplyr `select` Function**

```{r}
  litters_df
  select(litters_df, group, litter_number)  # select column group & litter_number
  select(litters_df, group, gd0_weight:gd_of_birth)  # select column group & all groups from gd0_weight to gd_of_birth
  select(litters_df, -litter_number)  # get rid of column "litter_number" and keep everything else
  select(litters_df, starts_with("gd"))    # a quick way to select columns by matching string patterns
  select(litters_df, litter_number, everything())   # select "litter_number" column and then all else columns
```

#### **Rename Columns**

```{r}
  select(litters_df, GROUP=group, LITTER_NUMBER = litter_number)    # rename column group and litter_number while extracted
  rename(litters_df, GROUP=group, LITTER_NUMBER = litter_number)    # rename column group and litter_number and display all dataset
```

#### **Relocate Column when Displaying (but not changing the original column sequence)**
```{r}
  relocate(litters_df, litter_number)
  litters_df
```

#### **Dplyr `Filter` Function**
```{r}
  filter(litters_df, gd0_weight < 22)
  filter(litters_df, gd0_weight >= 22)
  filter(litters_df, gd0_weight == 27)
  filter(litters_df, gd0_weight != 27)
  filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
  filter(litters_df, group == "Mod8")
  filter(litters_df, group %in% c("Con7", "Mod8"))
```

#### **Dplyr `mutate` Function**
*The mutate() function is part of the dplyr package. It is used to create new columns or modify existing columns in a data frame without changing the structure of the data frame*
```{r, message=FALSE, warning=FALSE}
  litters_mutated = mutate(litters_df, wt_gain=as.numeric(gd18_weight)-as.numeric(gd0_weight), new_col=str_to_lower(group) ) 
  # str_to_lower() converts all string arguments inside to lower case
  litters_mutated
```

#### **Dplyr `arrange` Function**

```{r}
  arrange(litters_df, pups_born_alive)        # order the pops_born_alive column in ascending order
  arrange(litters_df, desc(pups_born_alive))  # or you can use the optional desc() to sort in descending order
```

#### **`%>%`**
```{r, results="HIDE", message=FALSE}
  # originally, we would typically do
  litters_data_raw = read_csv("../../Datasets/FAS_litters.csv")
  litters_clean_name = janitor::clean_names(litters_data_raw)
  littesr_data_selected = select(litters_clean_name, -pups_born_alive)
  litters_without_missing = drop_na(litters_mutated, gd0_weight)
```

*NOW USE THE PIPE OPERATOR INSTEAD*
##### `%>%` can be interpreted as "**AND THEN**"
```{r}
litters_df = 
    read_csv("../../Datasets/FAS_litters.csv") %>%
    janitor::clean_names() %>%
    select(-pups_born_alive) %>%
    drop_na(gd0_weight)
litters_df
```

